Persistence in functional analysis and a correction to a theorem of Morse

### 2021-06-15 inequalities.tex
version with stability

\section{Morse Inequalities} \label{s:inequalities}

We now present a description of Morse inequalities from the viewpoint of persistence theory which, when applied in the context of Functional Topology, agrees with Morse's original definitions.

Let $M$ be a graded q-tame persistence module, and let $\mathfrak{m}_d$ be the multiplicity function of the barcode of the $d$-dimensional part of $M$, having values in the extended natural numbers.

To keep track of the number of $d$-dimensional homological events at filtration value $t$ that persist for at least time $e$, Morse defined the $(d, t, e)$-\textit{cap numbers}, which, when the associated persistence module has a barcode and is upper semi-continuous\footnote{A persistence module $M$ is \emph{upper semi-continuous} if $M_{t} \to \lim_{u > t} M_{u}$ is an isomorphism for all~$t$.}, correspond to the number of bars with length greater than $e$ in the $d$-dimensional barcode with left endpoint $t$, plus the number of bars with length greater than $e$ in the $(d-1)$-dimensional barcode with right endpoint $t$ (killed by something $d$-dimensional).
Explicitly,
\[
    m_{d}^{> e}(t) =
    \sum_{\substack{p \in \R \\ t-p>e}} \mathfrak{m}_{d-1}(p, t) +
    \sum_{\substack{q \in \R \\ q-t>e}} \mathfrak{m}_d(t, q).
\]

If $m_{d}^{> e}(t)$ is non-zero for only finitely many $t$, we obtain the $(d,e)$-cap numbers 
\[
m_{d}^{> e } = \sum_{t} m_{d}^{> e}(t) = \sum_{\substack{(p,q)\\q-p>e\\q\neq\infty}}\mathfrak{m}_{d-1}(p,q)+\sum_{\substack{(p,q)\\q-p>e\\p\neq-\infty}}\mathfrak{m}_d(p,q).
\] 
One can show that this finiteness assumption is satisfied if $M$ is initially and eventually constant, i.e., if there exist $t_0$ and $t_1$ with $M_s = M_{t_0}$ for all $s \leq t_0$ and $M_u = M_{t_1}$ for all $u \geq t_1$. In particular, this is the case for the persistent homology of the sublevel set filtration of a bounded function.

If $X$ is a closed manifold and $F$ is a Morse function, then the cap numbers $m_{d}^{> e}(F)$ associated to the persistent homology of the sublevel set filtration of $F$ agree with the number of critical points of $F$ with index $d$ for all sufficiently small $e > 0$. As an analogue to the Betti numbers of $X$ appearing in the usual Morse inequalities, Morse defines quantities $p_{d}$, which, under the same assumptions as before, can be expressed in the language of persistence as
\[
p_{d} = \sum_{p} \mathfrak{m}_d(p,\infty).
\]

Provided that $\mathfrak{m}_d(-\infty, p) = 0$ for all $p$, an easy calculation shows that in this language we have Morse inequalities for any graded q-tame persistence module given by
\begin{equation*}
    \sum_{i=1}^d (-1)^i (m^{>e}_{d-i} - p_{d-i}) \ \geq\  0
\end{equation*}
for any dimension $d$. As a special case, this yields Morse inequalities for any bounded real-valued function $F$ whose sublevel set filtration has q-tame and upper semi-continuous persistent homology. As we have discussed previously, these algebraic conditions are satisfied in the setting of Functional Topology, where \v{C}ech homology is used and the sublevel set filtration is assumed to be compact and $\HLC$.

>>>>

\section{Generalized Morse inequalities from persistence diagrams} \label{s:inequalities}

In this section we prove that $q$-tame persistence modules satisfy a general version of Morse inequalities that specializes to the usual Morse inequalities in the smooth context, as well as to the version used by Morse--Tompkins to prove their Unstable Minimal Surface Theorem, see \cref{s:surfaces} for more details on this result.
We deduce this general version from the existence of persistence diagrams, \cref{thm:q-tame modules have barcodes}.

Fix a $\Z$-graded q-tame persistence module $M$.
For simplicity, the reader may think of $M$ as the persistent homology of a q-tame filtration.
By \cref{thm:q-tame modules have barcodes} $M$ has a persistence diagram in every degree $d$ with multiplicity function denoted $\mathfrak{m}_d \colon \multiplicityDomain \to \N$.

To keep track of the number of $d$-dimensional homological events at filtration value $t$ that persist for at least time $e$ but not indefinitely, \citet{Morse.1940} defined the $(d, t, e)$-\textit{cap numbers} of a filtration, which, when the associated persistence module has a barcode and is upper semi-continuous\footnote{A persistence module $M$ is \emph{upper semi-continuous} if $M_{t} \to \lim_{u > t} M_{u}$ is an isomorphism for all~$t$.}, correspond to the number of finite bars with length greater than $e$ in the $d$-dimensional barcode with left endpoint $t$, plus the number of finite bars with length greater than $e$ in the $(d-1)$-dimensional barcode with right endpoint $t$ (killed by something $d$-dimensional).

In analogy, we may define the $(d, t, e)$-\textit{cap number} of our graded q-tame persistence module $M$ in terms of its persistence diagram as
\[
m_{d}^{> e}(t) =
\sum_{\substack{p \in \R \\ t - p > e}} \mathfrak{m}_{d-1}(p, t) +
\sum_{\substack{q \in \R \\ q - t > e}} \mathfrak{m}_d(t, q).
\]
Note that finiteness of these quantities is ensured by the q-tameness of $M$.
% which are well-defined natural numbers because of the q-tameness of $M$.

Whenever the sums below are well-defined, we also consider the \emph{$(d,e)$-cap numbers} 
\[
m_{d}^{> e } 
= \sum_{t} m_{d}^{> e}(t) 
=
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > e \\ q \neq \infty}} \mathfrak{m}_{d-1}(p,q)
+
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > e \\ p \neq -\infty}}\mathfrak{m}_d(p,q).
\] 
They are well-defined for example if $M$ is the persistent homology of the sublevel set filtration of a bounded function. This can be deduced from the following general result.

\begin{thm}
    Assume that $N$ is a q-tame persistence module that is initially and eventually constant, i.e., that there exist $t_0$ and $t_1$ with $N_s = N_{t_0}$ for all $s \leq t_0$ and $N_u = N_{t_1}$ for all $u \geq t_1$, and let $\mathfrak{n}$ be the multiplicity function of the persistence diagram of $N$. Then for each $e>0$ we have 
    \[
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > e } } \mathfrak{n} (p,q) < \infty.
    \]
\end{thm}

\begin{proof}
    We split the sum whose finiteness we want to show in two parts
    \[
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > e } } \mathfrak{n} (p,q)
    =
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > e } } \mathfrak{n} (p,q)
    +
    \sum_{ \substack{ (p,q) \in T \\ q-p > e } } \mathfrak{n} (p,q)
    \]
    and show for each summand separately that it is finite. 
    Here, $T$ denotes the triangle
    \[
    T = \{(p,q) \in \multiplicityDomain \mid t_0 \leq p < q \leq t_1\}.
    \]

    For the first summand, observe that because $N$ is constant below $t_0$ and constant above $t_1$, we have $\mathfrak{n}(p,q) = 0$ whenever one of $-\infty < p < t_0$, $q < t_0$, $p > t_1$ or $t_1 < q < \infty$ holds.
    This implies 
    \[
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > e } } \mathfrak{n} (p,q)
    =
    \sum_{t_0 < q < t_1} \mathfrak{n} (-\infty,q)
    +
    \sum_{t_0 < p < t_1} \mathfrak{n} (p, \infty)
    \]
    which is clearly finite because $N$ is q-tame.

    For the second summand, note that  
    \[
    \sum_{ \substack{ (p,q) \in T \\ q-p > e } } \mathfrak{n} (p,q)
    \leq
    \sum_{(p,q) \in T^{\geq e}} \mathfrak{n} (p,q),
    \]
    where $T^{\geq e}$ is the smaller triangle 
    \[
    T^{\geq e} = \{(p,q) \in T \mid q-p \geq e\}.
    \]
    Thus, in order to prove the theorem, it suffices to show that we have 
    \[
    \sum_{(p,q) \in T^{\geq e}} \mathfrak{n} (p,q) 
    < 
    \infty.
    \]
    To do this, we consider open quadrants 
    \[
    Q(x, y) = \{ (p, q) \in \R^2 \mid p < x \text{ and } y < q \}.
    \]
    Covering the compact set $T^{\geq e}$ by the open quadrants $Q \left(x, x + \frac{e}{2} \right)$ for $x \in \R$, we may choose a finite subcover given by, say, $x_1,\dots, x_n$. We obtain
    \[
    \sum_{(p,q) \in T^{\geq e}} \mathfrak{n} (p,q) 
    \ \leq \
    \sum_{i=1}^n \sum_{\substack{(p, q) \in \\ Q (x_i, x_i + \frac{e}{2})}} \mathfrak{n}(p,q). 
    \]
    Each of sums $\sum_{(p,q) \in Q \left(x_i, x_i + \frac{e}{2} \right)} \mathfrak{n}(p,q)$ over the quadrants $Q \left(x_i, x_i + \frac{e}{2} \right)$ is finite since $N$ is q-tame (which is where the name q-tame or \emph{quadrant}-tame comes from), see \cite[Section 3.8]{Chazal.2016a}. This finishes the proof.
\end{proof}

%If $X$ is a closed manifold and $F$ is a Morse function, then the cap numbers $m_{d}^{> e}(F)$ associated to the persistent homology of the sublevel set filtration of $F$ agree with the number of critical points of $F$ with index $d$ for all sufficiently small $e > 0$.

Comparing to the usual Morse inequalities, the cap numbers in dimension $d$ act like the number of critical points with index $d$. As an analogue to the Betti numbers of the manifold appearing in the usual Morse inequalities, Morse defines quantities $p_{d}$, which, under the same assumptions as before, can be expressed in the language of persistence as
\[
p_{d} \ = \!\! \sum_{p \in \R \cup \{-\infty\}} \mathfrak{m}_d(p,\infty),
\]
which is also the dimension of the colimit of the degree $d$ part of $M$.
%Because $M$ is 0-interleaved with $\bigoplus_(p,q) \bigoplus_{\mathfrak{m}(p,q)} C((p,q))$ and colimits commute with direct sums and are invariant up to finite interleaving distance.

\begin{thm}
    Let $M$ be a graded q-tame persistence module with $m_{d}^{> e }$ and $p_{d}$ finite for all $d$ and $e$. If $\mathfrak{m}_d(-\infty, p) = 0$ for all $p \in \R \cup \{\infty\}$ and all $d$, then we have Morse inequalities
    \begin{equation} \label{e:morse inequalities}
        \sum_{d=0}^n \ (-1)^{n-d} (m^{>e}_{d} - p_{d}) \ \geq\  0
    \end{equation}
    for any dimension $d$ and any $e > 0$.
\end{thm}

\begin{proof}
    Using a simple argument carried out for example in \cite[Section 11]{Morse.1940}, the claimed inequalities are equivalent to the existence of a sequence $(\nu_d)_d$ of non-negative integers with $m^{>e}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$.

    Since we assume $\mathfrak{m}_d(-\infty, p) = 0$ for all $p$, we have
    \begin{align*}
    m^{>e}_{d}
    \ &=
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > e \\ q \neq \infty}} \mathfrak{m}_{d-1}(p,q)
    \ +
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > e \\ p \neq -\infty}}\mathfrak{m}_d(p,q)
    \\
    &=
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > e }} \mathfrak{m}_{d-1}(p,q)
    \ +
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > e \\ p \neq -\infty}}\mathfrak{m}_d(p,q)
    \end{align*}
    and
    \[
    p_{d}
    \ = \!\!
    \sum_{p \in \R \cup \{-\infty\}} \mathfrak{m}_d(p,\infty)
    \ = \
    \sum_{p \in \R} \mathfrak{m}_d(p,\infty),
    \]
    which yields
    \begin{align*}
    m^{>e}_{d} - p_{d}
    &= 
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > e }} \mathfrak{m}_{d-1}(p,q)
    +
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > e \\ p \neq -\infty}}\mathfrak{m}_d(p,q)
    -
    \sum_{p \in \R} \mathfrak{m}_d(p,\infty)
    \\
    &=
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > e }} \mathfrak{m}_{d-1}(p,q)
    +
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > e }}\mathfrak{m}_d(p,q),
    \end{align*}
    so the sequence given by $\nu_d = \sum_{\substack{(p,q) \in \R^2 \\ q - p > e }}\mathfrak{m}_d(p,q)$ satisfies $m^{>e}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$. This finishes the proof.
\end{proof}
As we have remarked, the finiteness assumptions are satisfied if $M$ is initially and eventually constant, so as a special case, the theorem yields Morse inequalities for any bounded real-valued function whose sublevel set filtration has q-tame persistent homology. This includes classical Morse functions on closed manifolds, for which the inequalities in our theorem agree with the classical Morse inequalities if $e > 0$ is chosen sufficiently small.

Our inequalities, as presented in \eqref{e:morse inequalities}, also agree with Morse's historical version from functional topology if the persistent homology of the sublevel set filtration is not only q-tame but also upper semi-continuous, which is satisfied for compact sublevel set filtrations if we use \v{C}ech homology.

### 2021-06-15 introduction.tex
version with stability

The interplay between the critical set of a function and the topology of its domain is a cornerstone of modern mathematics.
Nowadays, when thinking about the pioneering work of Marston Morse, our first association is possibly with a smooth function on a closed differential manifold, but, as we describe next, that should not be our the last.
Morse theory in the smooth context was masterfully presented in Milnor's famous book on the subject \cite{MR0163331}, where he also focused on applications beyond this restricted setting;
giving a new proof of Bott's periodicity by applying Morse theory to the energy functional of paths in a Riemannian manifold.
Another important example of the use of Morse's insights in an infinite setting is Floer's work on the Arnold conjecture and its many ramifications in symplectic topology, as surveyed for example in \cite{MR1702944}.
Morse himself worked in a very general setting, publishing in the 1930s a series of papers \cite{Morse.1937, Morse.1938, Morse.1940, MR9102} in which he established the key results of Morse theory in the broad context defined by semi-continuous functionals on metric spaces.
He named this body of work functional topology and used it to study minimal surfaces motivated by Douglas' solution to Plateau’s Problem \cite{Douglas.1931}.
In particular, Morse and Tompkins \cite{Morse.1939, Morse.1941} used these techniques to prove their Unstable Minimal Surface Theorem which gives conditions for the existence of critical points of the Douglas functional that do not locally minimize area.

While Morse's work on functional topology did not have a long lasting impact in minimal surface theory, many of his ideas have been rediscovered in the study of persistent homology.
A method originated in applied topology and based on the functoriality of the homology construction, that provides robust and effectively computable invariants of filtered spaces.
The successes of this technique have motivated the development of an abstract and rich theory of persistence, lying in the intersection of representation theory, geometry, and topology, and whose influence extends beyond it original domain.

One of the primary reasons for the use of persistence theory in geometry is the fact that its basic invariants, known as barcodes, can be organized into a metric space, and that passing from filtrations to barcodes is $1$-Lipschitz.
This often allows to recast geometric questions about general filtered spaces into a combinatorial and typically finite metric model.

For the existence of barcodes some form of finite-dimensionality must be assumed of the  filtered space being considered.
The most general condition associated to this problem is known as quadrant-tameness or q-tameness for short \cite{bibid}.
It simply states that the inclusion of distinct sublevel sets induces in homology with field coefficients a linear map with finite dimensional rank.

Even though Morse did not have a structural description of the homology of a filtration like the one provided by barcodes, he emphasized many aspects that are crucial for their existence, including the use of field coefficients and the restriction to q-tame filtrations.
These allowed him to state and prove a version his famous inequalities for general filtered spaces using homotopical critical points.
These inequalities are crucial for applications and we reinterpret them in the modern language of barcodes in \cref{s:inequalities}.

Since general applications may take us away from spaces with the homotopy type of CW-complexes, the choice of homology construction is an important element of the discussion.
Two types of properties of these constructions are important for the existence of barcodes. On one hand, those ensuring that local-connectivity properties of sublevel set filtrations are sufficient for q-tameness which, thanks to our analysis, boil down to the existence of Mayer-Vietories exact sequences.
On the other, those ensuring that q-tame persistence modules admit a multiplicative interval decomposition, which was shown in \cite{schmahl2020structure} to follow from the preservation of certain categorical limits.

Our analysis allows us to show the insufficiency of the local-connectivity assumption used by Morse and Tompkins to state that the filtration defined by Douglas's functional is q-tame.
This is because it falls into the class ruled out by \cref{thm:counterexample}.
The existence of Morse inequalities requires q-tameness and, hence, so does the Unstable Minimal Surface Theorem of Morse and Tompkins.
Nonetheless, this filtration satisfies a stronger condition which \cref{t:strong local connectedness implies q-tameness} ensures to be sufficient

By reinterpreting the work of Morse in the modern context of persistence theory, and clarifying the conditions needed to prove the existence of barcodes for a broad class of sublevel set filtrations, we aim to increase the number and depth of interactions occurring between persistence theory on one side, and geometry, topology, and functional analysis on the other.

>>>>

The interplay between the critical set of a function and the topology of its domain is a cornerstone of modern mathematics.
Nowadays, when thinking about the pioneering work of Marston Morse, our first association is possibly with a smooth function on a closed differential manifold, but, as we describe next, that should not be our the last.
Morse theory in the smooth context was masterfully presented in Milnor's famous book on the subject \cite{MR0163331}, where he also focused on applications beyond this restricted setting;
giving a new proof of Bott's periodicity by applying Morse theory to the energy functional of paths in a Riemannian manifold.
Another important example of the use of Morse's insights in an infinite setting is Floer's work on the Arnold conjecture and its many ramifications in symplectic topology, as surveyed for example in \cite{MR1702944}.
Morse himself worked in a very general setting, publishing in the 1930s a series of papers \cite{Morse.1937, Morse.1938, Morse.1940, MR9102} in which he established the key results of Morse theory in the broad context defined by semi-continuous functionals on metric spaces.
He named this body of work functional topology and used it to study minimal surfaces motivated by Douglas' solution to Plateau’s Problem \cite{Douglas.1931}.
In particular, Morse and Tompkins \cite{Morse.1939, Morse.1941} used these techniques to prove their Unstable Minimal Surface Theorem giving conditions for the existence of critical points of the Douglas functional that do not locally minimize area.

While Morse's work on functional topology did not have a long lasting impact in minimal surface theory, many of his ideas have flourished in other domains.
% Many of his ideas, specially when applied to the finite dimensional case, have been rediscovered in the study of persistent homology.
% , a method originated in applied topology and based on applying a homology functor to a filtered space,
We now review one such area.
The technique known as persistence homology provides robust and effectively computable invariants of filtered spaces through the application of a homology functor.
Its successes in applied topology have motivated the development of a refined abstract theory of persistence, lying in the intersection of representation theory, geometry, and topology.
The influence of this theory and its sophistication have seen rapid growth in recent years, motivated in part by the fact that
% has been growing and the and whose influence extends beyond finite dimensional settings.
% One of the primary reasons being that
its basic invariants, known as barcodes, can be organized into a metric space, and that, when they exists, the passage from filtrations to barcodes is Lipschitz.
This often allows to recast geometric questions about general filtered spaces into a combinatorial and typically finite metric model.

In this work we focus on the existence of barcode decompositions for a broad class of filtered spaces arising from the sublevel sets of real valued functions.
One important consequence we present is the existence of Morse inequalities when such decompositions are available.

For general persistence modules, a well studied condition for the existence of a barcode decomposition is q-tameness \cite{Chazal.2016a,Chazal.2016b}.
It simply states that all non-identity linear maps in the persistence module have finite dimensional rank.
In order to obtain the q-tameness of the persistence module associated to a sublevel set filtration, we study different notions of local-connectivity.
Showing their sufficiency and limitations in different context.

Having conditions for general sublevel set filtration that ensure the existence of barcodes and Morse inequalities is important to extend the influence of persistence methods into infinite dimensional geometrical settings, like the one originally considered by Morse and Topmkins.
As a consequence of our analysis we identify and fill a gap in the proof the Unstable Minimal Surface Theorem.

% Even though Morse did not have a structural description of the homology of a filtration like the one provided by barcodes, he emphasized many aspects that are crucial for their existence, including the use of field coefficients and the restriction to q-tame filtrations.
% These allowed him to state and prove a version his famous inequalities for general filtered spaces using homotopical critical points.
% These inequalities are crucial for applications and we reinterpret them in the modern language of barcodes in \cref{s:inequalities}.

% Since general applications may take us away from spaces with the homotopy type of CW-complexes, the choice of homology construction is an important element of the discussion.
% Two types of properties of these constructions are important for the existence of barcodes.
% On one hand, those ensuring that local-connectivity properties of sublevel set filtrations are sufficient for q-tameness which, thanks to our analysis, boil down to the existence of Mayer-Vietories exact sequences.
% On the other, those ensuring that q-tame persistence modules admit a multiplicative interval decomposition, which was shown in \cite{schmahl2020structure} to follow from the preservation of certain categorical limits.

% Our analysis allows us to show the insufficiency of the local-connectivity assumption used by Morse and Tompkins to state that the filtration defined by Douglas's functional is q-tame.
% This is because it falls into the class ruled out by \cref{thm:counterexample}.
% The existence of Morse inequalities requires q-tameness and, hence, so does the Unstable Minimal Surface Theorem of Morse and Tompkins.
% Nonetheless, this filtration satisfies a stronger condition which \cref{t:strong local connectedness implies q-tameness} ensures to be sufficient

% By reinterpreting the work of Morse in the modern context of persistence theory, and clarifying the conditions needed to prove the existence of barcodes for a broad class of sublevel set filtrations, we aim to increase the number and depth of interactions occurring between persistence theory on one side, and geometry, topology, and functional analysis on the other.

### 2021-06-15 main.tex
version with stability

\immediate\write18{bibtex \jobname}
\documentclass{amsart}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{tikz-cd}
\usepackage[numbers,sort,compress]{natbib}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[bookmarks=true,
bookmarksnumbered=true, breaklinks=true,
pdfstartview=FitH, hyperfigures=true,
plainpages=false, naturalnames=true,
colorlinks=true, pagebackref=true,
pdfpagelabels]{hyperref}
\hypersetup{
	linktocpage,
	colorlinks,
	citecolor=blue,
	linkcolor=blue,
	urlcolor=blue}
\usepackage[capitalize, noabbrev]{cleveref}

@@

% commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\HT}{\mathrm{HT}}
\newcommand{\HLC}{\mathrm{HLC}}
\newcommand{\piT}{\pi\mathrm{T}}
\newcommand{\LC}{\mathrm{LC}}
\newcommand{\NH}{\mathrm{NH}}
\newcommand{\piLC}{\pi\mathrm{LC}}
\newcommand{\HE}{\mathbb{H}^d}
\renewcommand{\H}{\widetilde{H}}
\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\im}{im}
\newcommand\CH{\check{H}}
\newcommand{\Mod}[1]{#1-\mathbf{Mod}}
\newcommand{\PFD}{\mathrm{PFD}}
\newcommand{\PFG}{\mathrm{PFG}}
\DeclareMathOperator{\Nrv}{Nrv}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Vietoris}{Vtr}
\DeclareMathOperator{\Balls}{Balls}
\DeclareMathOperator*{\colim}{colim}
\DeclareMathOperator{\crit}{crit}
\newcommand\MCH{M\check{H}}
\newcommand\VH{VH}
\newcommand\MVH{MVH}

% margins
\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-2.5in}
\setlength{\textheight}{\paperheight}
\addtolength{\textheight}{-2.5in}
\calclayout

@@

\title[Persistence in functional topology]{Persistence in functional topology: local-connectivity, q-tameness and a correction to a theorem of Morse}

@@

\begin{abstract}
	During the 1930's Morse developed the theory of functional topology relating the critical set of a semi-continuous functional and the topology of its sublevel sets.
	Morse and Tompkins applied this body of work in the setting defined by Douglas' solution to Plateau's Problem, and proved with it their Unstable Minimal Surface Theorem.
	Many of the key insights of functional topology have been rediscovered in the study of persistent homology, with the notion of q-tameness being as central to functional topology as it is to persistence theory.
	With applications to geometry in mind, in this article we present a modern treatment of the relationship between local-connectivity properties of sublevel set filtrations, Morse inequalities, and the q-tameness of their associated persistence modules.
	As a consequence of our analysis we identify a gap in the proof of the Unstable Minimal Surface Theorem of Morse and Tompkins and provide a correction.
\end{abstract}

\maketitle
\tableofcontents
\input{introduction}
\input{homology}
\input{persistence}
\input{inequalities}
\input{connectivity}
\input{surfaces}
\appendix
\input{vietoris}

>>>>

\immediate\write18{bibtex \jobname}
\documentclass{amsart}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{tikz-cd}
\usepackage[numbers,sort,compress]{natbib}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[bookmarks=true,
bookmarksnumbered=true, breaklinks=true,
pdfstartview=FitH, hyperfigures=true,
plainpages=false, naturalnames=true,
colorlinks=true, pagebackref=true,
pdfpagelabels]{hyperref}
\hypersetup{
	linktocpage,
	colorlinks,
	citecolor=blue,
	linkcolor=blue,
	urlcolor=blue}
\usepackage[capitalize, noabbrev]{cleveref}
\usepackage{todo}

@@

% commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\HT}{\mathrm{HT}}
\newcommand{\HLC}{\mathrm{HLC}}
\newcommand{\piT}{\pi\mathrm{T}}
\newcommand{\LC}{\mathrm{LC}}
\newcommand{\NH}{\mathrm{NH}}
\newcommand{\piLC}{\pi\mathrm{LC}}
\newcommand{\HE}{\mathbb{H}^d}
\renewcommand{\H}{\widetilde{H}}
\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\rad}{rad}
\newcommand\CH{\check{H}}
\newcommand{\Mod}[1]{#1-\mathbf{Mod}}
\newcommand{\PFD}{\mathrm{PFD}}
\newcommand{\PFG}{\mathrm{PFG}}
\DeclareMathOperator{\Nrv}{Nrv}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Vietoris}{Vtr}
\DeclareMathOperator{\Balls}{Balls}
\DeclareMathOperator*{\colim}{colim}
\DeclareMathOperator{\crit}{crit}
\newcommand\MCH{M\check{H}}
\newcommand\VH{VH}
\newcommand\MVH{MVH}
\newcommand{\obs}{\dashrightarrow}
\newcommand{\multiplicityDomain}{\mathcal{E}}
\newcommand{\Top}{\mathsf{Top}}
\newcommand{\Vect}{\mathsf{Vect}}
\newcommand{\anibal}[1]{\textcolor{blue}{\underline{Anibal}: #1}}

% margins
\setlength{\textwidth}{\paperwidth}
\addtolength{\textwidth}{-2.3in}
\setlength{\textheight}{\paperheight}
\addtolength{\textheight}{-2.3in}
\calclayout

@@

\title[Persistence in functional analysis]{Persistence in functional analysis and a correction to a theorem of Morse}

@@

\begin{abstract}
	During the 1930's Morse developed the theory of functional topology relating the critical set of a semi-continuous functional and the topology of its sublevel sets.
	Morse and Tompkins applied this body of work in the setting defined by Douglas' solution to Plateau's Problem, and proved with it their Unstable Minimal Surface Theorem.
	Many of the key insights of functional topology have been rediscovered in the study of persistent homology, with the notion of q-tameness being as central to functional topology as it is to persistence theory.
	With applications to geometry in mind, we present a modern treatment of local-connectivity of sublevel set filtrations, q-tameness of their associated persistence homology, and Morse inequalities.
	As a consequence of our analysis we identify and correct a gap in the proof of the Unstable Minimal Surface Theorem of Morse and Tompkins.
\end{abstract}

\maketitle
\tableofcontents
\input{introduction}
\input{persistence}
\input{inequalities}
\input{connectivity}
\input{surfaces}
\appendix
\input{vietoris}
